Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-19 paper.tex
save

% For figures
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{amsmath}

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two main components: (i) a
linkage function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
linkage function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
a small set of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a linkage function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific linkage functions tailored for author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}. These later
approaches have the advantage of not requiring  hand-labeled data, but often do
not perform as well as the supervised approaches. To reconcile both worlds,
semi-supervised methods make use of small, manually verified, clusters of
publications and/or of high-precision domain-specific rules to build a training
set of pairs of publications, from which a linakge function is then built
using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.
Coincidentally, semi-supervised approaches also allow for the tuning of the
clustering algorithm when the later is applied to a mixed set of labeled
and unlabeled publications, e.g., by maximizing some clustering performance
metric on the known clusters \citep{levin2012citation}.

@@

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures, for
later reuse during clustering to group all signatures from a same author.
Formally, the goal is to build a function $d: {\cal S} \times {\cal S} \mapsto
[0, 1]$, such that $d(s_1, s_2)$ tends to $0$ if both signatures $s_1$ and
$s_2$ belong to the same author, and $1$ otherwise.

@@

\textit{Features engineering.}

>>>>

% For figures
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}

% Maths
\usepackage{amsmath}
\usepackage{amsfonts}

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two main components: (i) a
linkage function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
linkage function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
a small set of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a linkage function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific, often manually designed, linkage functions tailored for author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}. These later
approaches have the advantage of not requiring  hand-labeled data, but often do
not perform as well as the supervised approaches. To reconcile both worlds,
semi-supervised methods make use of small, manually verified, clusters of
publications and/or of high-precision domain-specific rules to build a training
set of pairs of publications, from which a linakge function is then built
using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.
Coincidentally, semi-supervised approaches also allow for the tuning of the
clustering algorithm when the later is applied to a mixed set of labeled
and unlabeled publications, e.g., by maximizing some clustering performance
metric on the known clusters \citep{levin2012citation}.

@@

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures, for
later use during clustering to group all signatures from a same author.
Formally, the goal is to build a function $d: {\cal S} \times {\cal S} \mapsto
[0, 1]$, such that $d(s_1, s_2)$ tends to $0$ if both signatures $s_1$ and
$s_2$ belong to the same author, and $1$ otherwise.

@@

\textit{Input features.} In most cases, supervised learning algorithms assume
the input space ${\cal X}$ to be numeric (e.g., $\mathbb{R}^p$), making them
not directly applicable to structured input spaces such as ${\cal S} \times
{\cal S}$. Following previous works, pairs of signatures $(s_1, s_2)$ are
transformed to vectors $v \in \mathbb{R}^p$ by building so-called similarity
profiles \citep{treeratpituk2009disambiguating} on which supervised learning is
then carried on. In this work, we design and evaluate 15 standard input features based
on the comparison of signature fields, as reported in Table~\ref{table:features}.
As an illustrative example, the \textit{Full name} feature corresponds to ...

\begin{table}
\caption{Input features}
\label{table:features}
\centering
\begin{tabular}{|l|l|}
  \hline
  \textbf{Feature} & \textbf{Comparator}\\
  \hline
  \hline
  Full name & Cosine similarity of TF-IDF\\
  Given names & Cosine similarity of TF-IDF\\
  First given name & Jaro-Winkler distance\\
  Second given name & Jaro-Winkler distance\\
  Given name initial & Equality\\
  Affiliation & Cosine similarity of TF-IDF\\
  Co-authors & Cosine similarity of TF-IDF\\
  Title & Cosine similarity of TF-IDF\\
  Journal & Cosine similarity of TF-IDF\\
  Abstract & Cosine similarity of TF-IDF\\
  Keywords & Cosine similarity of TF-IDF\\
  Collaborations & Cosine similarity of TF-IDF\\
  References & Cosine similarity of TF-IDF\\
  Subject & Cosine similarity of TF-IDF\\
  Year difference & Absolute difference\\
  \hline
\end{tabular}
\end{table}

% ethnicity features
